### 2024-06-16 thesis/ms-fletcher-porter.tex
Work on the Solid Mechanics case study

The industry partner's requested change occurred in the IO (input/output) layer. The old system had an inverter that drives a servomotor, a safety valve to enable the system, and an array of pressure, temperature, and flow sensors to observe the state of the hydraulic system. The inverter and the pressure and temperature sensors all communicated over the CANopen protocol, the flow sensors gave analog inputs, and the safety valve was actuated with a digital signal. The CAN communication was interfaced by a Vector CAN/USB dongle~\cite{VectorCanUsbDongle}. All other IO went through a National Instruments data acquisition device (DAQ). In the new system, the forced change for the inverter to communicate over EtherCAT. This meant new IO hardware was needed to support this protocol, which was selected to be Beckhoff EtherCAT terminals. They provided the interface to connect the inverter, digital and analog IO, and a CAN bus for the sensor network~\cite{BeckhoffEL6751}. This made the DAQ unnecessary, but the CAN/USB dongle was still required due to deficiencies in Beckhoff's control interfaces, which are discussed with the control layer.

@@

Together, the stack for the old system was CANopen-based IO with a DAQ supplementing simple analog and digital signals, a workstation computer, Windows middleware with drivers and libraries to support communication with the CAN bus and DAQ, and a controller based on Matlab and Simulink. In the new system, the stack was a mix of EtherCAT and CANopen IO, a Beckhoff Industrial PC in concert with a workstation, PLC middleware with an application to initialize the CAN network, and a controller written in PLC languages with TwinCAT. This is illustrated in figure~\ref{fig:mechatronic_tech_stack_dolores}.

@@

In the IO layer, all three are similar. The six-legged and pipe robots have an array of on/off valves commanded by digital signals to pressurize their EFPAs. The servo valve has its position set with a small motor, which is controlled by a PWM signal. These signals are propagated through direct connection to pins on the computer, although Shinohara et al.\ note that there are circuits between stepping up power. The robots received command signals from a workstation over a serial connection. Hase et al.\ note that there was an intermediate chip on the six-legged robot to connect over USB. The servo valve computer had no reported inputs, though several sensors were connected to a data recording device.

@@

For the control layer, I wrote the code in the Sublime Text text editor~\cite{SublimeText} and configured it to automatically format the source code according to standards for the Rust language using the \verb|rustfmt| utility~\cite{rustfmt}.  Other text editors like Emacs, Vim, or VSCodium would have also been good choices, but Sublime is what is easiest for me.  This was done on a Windows workstation, but it was compiled and run on the Linux control computer.  To achieve this, I checked the changes on the workstation into a \verb|git| version control repository and pushed them to the control computer\footnote{A small issue that comes up is that \verb|git| generally doesn't allow users to push changes to a branch of a repository that the local user has checked out, which is the case here~\cite{GitConfigReceiveDenyCurrentBranch}.  This is to prevent data loss.  To enable this, \verb|git| can be configured to update the local branch only if there is no possibility of data loss by executing \verb|git config receive.denyCurrentBranch updateInstead|.}.  Then, though a remote terminal, I executed commands to compile and run the program.  This became quite annoying with how many steps there are, so I use the utility \verb|cargo-watch| to automatically execute the compile and run commands on the control computer when there are updates\footnote{The command was similar to \verb|cargo watch -x 'run -- eth0'|.}~\cite{CargoWatch}.  This still is somewhat tedious and so requires more development.  It also would have been mitigated by developing in some kind of Linux environment or by using a containerized environment like would be provided by Docker or Vagrant, which would have allowed running some tests on the development machine~\cite{DockerDev,VagrantDev}.  Most of the middleware was prepared with the Raspberry Pi OS image that was installed while comissioning the control computer, but there was some additional configuration, e.g. to install \verb|git|, that was done with the OS's \verb|apt| utility~\cite{apt}.  As well, library depndencies for the control program like Ethercrab were managed with the Rust \verb|cargo| utility~\cite{Cargo}.  The computer, once commissioned, did not require any development.  The IO layer's development was aided by a four-channel oscilliscope and a multimeter to validate electrical connections and ensure signal were propogating as expected.  It was also interesting to observe the timing jitter of the program, which was generally very small, on the order of microseconds.

@@

@@

This paper has introduced a theory of a mechatronic tech stack which describes function and tooling for mechatronic systems, namely as discrete layers of control, middleware, computer, and IO.  This theory is motivated by analogy to the web stack, used in IT software development, and to the OSI networking model, used to describe networked computer systems.  The OSI model also provided a means of discriminating the layers that make up the model, importantly by finding contrasts in the morphology, syntax, and semantics of the behavior of components of the control system.

>>>>

The industry partner's requested change occurred in the IO (input/output) layer. The old system had an inverter that drives a servomotor, a safety valve to enable the system, and an array of pressure, temperature, and flow sensors to observe the state of the hydraulic system. The inverter and the pressure and temperature sensors all communicated over the CANopen protocol, the flow sensors gave analog inputs, and the safety valve was actuated with a digital signal. The CAN communication was interfaced by a Vector CAN/USB dongle~\cite{VectorCanUsbDongle}. All other IO went through a National Instruments data acquisition device (DAQ). In the new system, the forced change for the inverter to communicate over EtherCAT.\ This meant new IO hardware was needed to support this protocol, which was selected to be Beckhoff EtherCAT terminals. They provided the interface to connect the inverter, digital and analog IO, and a CAN bus for the sensor network~\cite{BeckhoffEL6751}. This made the DAQ unnecessary, but the CAN/USB dongle was still required due to deficiencies in Beckhoff's control interfaces, which are discussed with the control layer.

@@

Together, the stack for the old system was CANopen-based IO with a DAQ supplementing simple analog and digital signals, a workstation computer, Windows middleware with drivers and libraries to support communication with the CAN bus and DAQ, and a controller based on Matlab and Simulink. In the new system, the stack was a mix of EtherCAT and CANopen IO, a Beckhoff Industrial PC in concert with a workstation, PLC middleware with an application to initialize the CAN network, and a controller written in PLC languages with TwinCAT.\ This is illustrated in figure~\ref{fig:mechatronic_tech_stack_dolores}.

@@

In the IO layer, all three are similar. The six-legged and pipe robots have an array of on/off valves commanded by digital signals to pressurize their EFPAs. The servo valve has its position set with a small motor, which is controlled by a PWM signal. These signals are propagated through direct connection to pins on the computer, although Shinohara et al.\ note that there are circuits between stepping up power. The robots received command signals from a workstation over a serial connection. Hase et al.\ note that there was an intermediate chip on the six-legged robot to connect over USB.\ The servo valve computer had no reported inputs, though several sensors were connected to a data recording device.

@@

For the control layer, I wrote the code in the Sublime Text text editor~\cite{SublimeText} and configured it to automatically format the source code according to standards for the Rust language using the \verb|rustfmt| utility~\cite{rustfmt}.  Other text editors like Emacs, Vim, or VSCodium would have also been good choices, but Sublime is what is easiest for me.  This was done on a Windows workstation, but it was compiled and run on the Linux control computer.  To achieve this, I checked the changes on the workstation into a \verb|git| version control repository and pushed them to the control computer\footnote{A small issue that comes up is that \verb|git| generally doesn't allow users to push changes to a branch of a repository that the local user has checked out, which is the case here~\cite{GitConfigReceiveDenyCurrentBranch}.  This is to prevent data loss.  To enable this, \verb|git| can be configured to update the local branch only if there is no possibility of data loss by executing \verb|git config receive.denyCurrentBranch updateInstead|.}.  Then, though a remote terminal, I executed commands to compile and run the program.  This became quite annoying with how many steps there are, so I use the utility \verb|cargo-watch| to automatically execute the compile and run commands on the control computer when there are updates\footnote{The command was similar to \verb|cargo watch -x 'run -- eth0'|.}~\cite{CargoWatch}.  This still is somewhat tedious and so requires more development.  It also would have been mitigated by developing in some kind of Linux environment or by using a containerized environment like would be provided by Docker or Vagrant, which would have allowed running some tests on the development machine~\cite{DockerDev,VagrantDev}.  Most of the middleware was prepared with the Raspberry Pi OS image that was installed while comissioning the control computer, but there was some additional configuration, e.g.\ to install \verb|git|, that was done with the OS's \verb|apt| package management utility~\cite{apt}.  As well, library depndencies for the control program like Ethercrab were managed with the Rust \verb|cargo| utility~\cite{Cargo}.  The computer, once commissioned, did not require any development.  The IO layer's development was aided by a four-channel oscilliscope and a multimeter to validate electrical connections and ensure signal were propogating as expected.  It was also interesting to observe the timing jitter of the program, which was generally very small, on the order of microseconds.

@@

The Solid Mechanics Laboratory at Aalto University conducts experiments to identify the properties of engineering materials under load particularly studying phenomena ranging from the micron to the meter scale.  As is typical for laboratories studying solid mechanics, they do this by subjecting test samples to a load schedule using a servo controlled hydraulic cylinder.

@@

This paper has introduced a theory of a mechatronic tech stack which describes function and tooling for mechatronic systems, namely as discrete layers of control, middleware, computer, and IO.\  This theory is motivated by analogy to the web stack, used in IT software development, and to the OSI networking model, used to describe networked computer systems.  The OSI model also provided a means of discriminating the layers that make up the model, importantly by finding contrasts in the morphology, syntax, and semantics of the behavior of components of the control system.
